\chapter{Introduction}

\section{\ee\ and \rtd}

Evidence presents the \ee\ RTOS, a multi-processor real-time operating 
system kernel, which is available for several platforms from Altera's 
NIOS II Softcores to the \dspic\ micro-controllers family. 

The \ee\ kernel introduces innovative concepts, mechanisms and programming
features to support the future generation of micro-controllers and multicore
systems-on-a-chip.

\ee\ offers the availability of a real-time scheduler and resource 
managers allowing the full exploitation of the power of new generation 
micro-controllers and multicore platforms while guaranteeing predictable 
real-time performance and retaining the programming model of 
conventional single processor architectures.

The advanced features provided by \ee\ are:

\begin{itemize}
\item Support for four conformance classes to match different
application requirements;
\item Support for preemptive and non-preemptive multitasking;
\item Support for fixed priority scheduling;
\item Support for stack sharing techniques, and one-shot task
model to reduce the overall stack usage;
\item Support for shared resources;
\item Support for periodic activations using Alarms;
\item Support for centralized Error Handling;
\item Support for hook functions before and after each context
switch;
\end{itemize}

The \ee\ kernel has been developed with the idea of providing the
minimal set of primitives which can beq used to implement a
multithreading environment. The \ee\ API as a reduced OSEK/VDX API,
providing support for thread activation, mutual exclusion, alarms, and 
counting semaphores.

The OSEK/VDX consortium provides the OIL language (OSEK Implementation 
Language) as a standard configuration language, which is used for the 
static definition of the RTOS objects which are instantiated and used 
by the application. \ee\ fully supports the OIL language for the 
configuration of real-time applications.

\ee\ is natively supported by \rtd, a tool suite for the automatic 
configuration and deployment of embedded applications which enables to 
easily exploit multi-processor architectures and achieve the desired 
performance without modifying the application source code. More details
about \rtd\ are available in a dedicated reference manual.

The purpose of this document is to describe in detail the \ee\
API. Details about specific architecture can be found in dedicated manuals,
available for download on the Evidence web site.



